---
title: "Day 24 Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
library(knitr)
library(dplyr)
library(stringr)
library(tidyverse)
library(readr)
library(collections)
options(scipen = 999)
```

```{r, eval=FALSE}
input1<-read_lines("Day24Sample1.txt")
input3<-read_lines("Day24Sample3.txt")

inplines<-which(str_detect(input,"inp"))
```



In today's version of "making searches out of things that shouldn't be searches
I tried to do this by hand (and eventually found a way to do it - thanks helpful internet post), but I kept getting lost - so, DFS our way through it, so it should be a little better than just trying to count down from 99,999,999,999.

First, build something that takes in an instruction list, an "input" and starting values for all of the registers and returns ending values for all the registers

```{r}
aluproc<-function(instructions,starter,startw,startx,starty,startz){
  regs<-dict()
  ### set the registers
  regs$set("w",startw)
  regs$set("x",startx)
  regs$set("y",starty)
  regs$set("z",startz)
  for(i in 1:length(instructions)){
    inst<-instructions[i]
    inst<-unlist(str_split(inst," "))
    a<-inst[2]
    if(length(inst)==3){
      ### change b to a number by getting it from the register or converting ig.
      b<-inst[3]
      if(b%in%letters){b<-regs$get(b)}else{b<-as.numeric(b)}}
    ### set up the functions
    switch(inst[1],
           "inp"={regs$set(a,starter)},
           "add"={regs$set(a,regs$get(a)+b)},
           "mul"={regs$set(a,regs$get(a)*b)},
           ### do not divide by 0
           "div"={if(b==0){return(NA)
           }else{regs$set(a,regs$get(a)%/%b)}},
           ### don't mod where the denominator is 0 or either is negative
           "mod"={if(b<=0||regs$get(a)<0){return(NA)
           }else{regs$set(a,regs$get(a)%%b)}},
           "eql"={if(regs$get(a)==b){regs$set(a,1)}else{regs$set(a,0)}},
           cat("***",inst[1],"\n","something went wrong\n"))}
  ### in something VERY SPECIFIC TO THIS DATA -  the answer won't work unless certain conditions are met - 
  ### if row 5 is div z 26, then end z MUST be less than start z
  if(instructions[5]=="div z 26"&& (regs$get("z")>startz)){return(NA)}
  return(c(regs$get("w"),regs$get("x"),regs$get("y"),regs$get("z")))}
```


So, how to make this into a highly specialized search

```{r}
findlowest<-function(allinstructions){
  s<-stack()
  ### this will count down from 9-1
  sapply(1:9,function(x){
    a<-list(list(c(x)),0,0,0,0)
    s$push(a)})
  beenthere<-dict()
  while(s$size()>0){
    now<-s$pop()
    module<-length(now[[1]])
    modrun<-aluproc(allinstructions[((module-1)*18+1):(module*18)],unlist(tail(now[[1]],1)),now[[2]],now[[3]],now[[4]],now[[5]])
    if(length(modrun)==1){next}
    if(modrun[4]>(26^3)){next}
    if(module<14){
      sapply(1:9,function(x){
        nxt<-list(c(now[[1]],x),modrun[1],modrun[2],modrun[3],modrun[4] )
        s$push(nxt)})
    }else{
      if(modrun[4]==0){return(list(now[[1]],modrun))}}}
  return("huh")}
  
```


leaving eval on false because it shows my data

```{r, eval=FALSE}
### with my input this takes 3.5 hours to run.  But it does run.
p1<-findlowest(input)
part1<-p1[[1]]
part1
part1<-as.numeric(str_flatten(part1))
part1
```


## Part 2

Do the opposite, 1-9 rather than 9-1

```{r}
findhighest<-function(allinstructions){
  s<-stack()
  #counts up from 1->9
  sapply(9:1,function(x){
    a<-list(list(c(x)),0,0,0,0)
    s$push(a)})
  beenthere<-dict()
  while(s$size()>0){
    now<-s$pop()
    module<-length(now[[1]])
    modrun<-aluproc(allinstructions[((module-1)*18+1):(module*18)],unlist(tail(now[[1]],1)),now[[2]],now[[3]],now[[4]],now[[5]])
    if(length(modrun)==1){next}
    if(modrun[4]>(26^3)){next}
    if(module<14){
      sapply(9:1,function(x){
        nxt<-list(c(now[[1]],x),modrun[1],modrun[2],modrun[3],modrun[4] )
        s$push(nxt)})
    }else{
      if(modrun[4]==0){return(list(now[[1]],modrun))}}}
  return("no solution\n")}
  
```

```{r, eval=FALSE}
p2<-findhighest(input)
part2<-p2[[1]]
part2<-as.numeric(str_flatten(part2))
part2
```

