---
title: "Day 18 Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
library(knitr)
library(dplyr)
library(stringr)
library(tidyverse)
library(readr)
library(collections)
options(scipen = 999)
```

```{r}
input<-read_lines("Day18Sample2.txt")
```

## Part 1
So, we'll need a bunch of functions

Add
```{r}
snailadd<-function(a,b){
  sa<-str_c("[",a,",",b,"]")
sa}
```

Snail reduce:
if there's an opportunity to explode or split, take it and repeat
```{r}
snailreduce<-function(a){
  ##while can either explode or reduce
  if(str_detect(a,"[0-9][0-9]+")){cansplit<-TRUE}else{cansplit<-FALSE}
  while(explosive(a)||cansplit){
    #### explode as much as possible
    while(explosive(a)){
      a<-snailexplode(a)
      #cat("explode to ",a,"\n")
      }
    ### split once if possible
    if(str_detect(a,"[0-9][0-9]+")){cansplit<-TRUE}else{cansplit<-FALSE}
    if(cansplit){a<-snailsplit(a)
    #cat("split to ",a,"\n")
    }
    #### find if there are more explosions and or splits
    if(str_detect(a,"[0-9][0-9]+")){cansplit<-TRUE}else{cansplit<-FALSE}}
a}
```


explosive:
checks to see if there's a pair that can explode

```{r}
explosive<-function(a){
  ### find all simple pairs
  prs<-str_locate_all(a,"\\[[0-9]+,[0-9]+\\]")
  y<-FALSE
  ### if there's at least one simple pair,
  if(length(prs)>0){y<-sapply(prs[[1]][,1],function(x){
    ### count the ['s & ]'s to the left
    l<-str_count(str_sub(a,1,x),"\\[")
    r<-str_count(str_sub(a,1,x),"\\]")
    ### if there are more than [ outweighs] by 4, something can explode
    exq<-if((l-r)>4){TRUE}else{FALSE}
    exq})}
  a<-(any(y))
  a}
```




Snail explode - explodes the right side, then the left side, then changes to 0.
```{r}
snailexplode<-function(a){
  ### find the first pair that can explode  (so yes, redo explosive, basically)
  prs<-str_locate_all(a,"\\[[0-9]+,[0-9]+\\]")
  y<-sapply(prs[[1]][,1],function(x){
    ### count the ['s & ]'s to the left
    l<-str_count(str_sub(a,1,x),"\\[")
    r<-str_count(str_sub(a,1,x),"\\]")
    ### if there are more than [ outweighs] by 4, something can explode
    exq<-if((l-r)>4){TRUE}else{FALSE}})
  ### its location
  explodemelocation<-prs[[1]][min(which(y)),]
  explodemestring<-str_sub(a,explodemelocation[1],explodemelocation[2])
  ### its values
  explodemevalue<-as.numeric(unlist(str_split(str_sub(a,explodemelocation[1]+1,explodemelocation[2]-1),",")))
  ## find the first number to the right
  rightnumval<-NA
  rightnumpos<-str_locate_all(str_sub(a,start=explodemelocation[2]),"[0-9]+")[[1]]
  if(nrow(rightnumpos)>0){rightnumpos<-rightnumpos[1,]+explodemelocation[2]-1
  rightnumval<-as.numeric(str_sub(a,rightnumpos[1],rightnumpos[2]))
  str_sub(a,rightnumpos[1],rightnumpos[2])<-rightnumval+explodemevalue[2]
  }
  ### find the first number to the left
  leftnumpos<-str_locate_all(str_sub(a,1,explodemelocation[1]),"[0-9]+")[[1]]
  ### find its position & value
  leftnumpos<-leftnumpos[nrow(leftnumpos),]
  leftnumval<-as.numeric(str_sub(a,leftnumpos[1],leftnumpos[2]))
  ### add the explode me value
  if(!is.na(leftnumval)){str_sub(a,leftnumpos[1],leftnumpos[2])<-leftnumval+explodemevalue[1]}
  ### replace with 0 - need to refind the explodeme
  prs<-str_locate_all(a,"\\[[0-9]+,[0-9]+\\]")
  y<-sapply(prs[[1]][,1],function(x){
    ### count the ['s & ]'s to the left
    l<-str_count(str_sub(a,1,x),"\\[")
    r<-str_count(str_sub(a,1,x),"\\]")
    ### if there are more than [ outweighs] by 4, something can explode
    exq<-if((l-r)>4){TRUE}else{FALSE}})
  ### its location
  explodemelocation<-prs[[1]][min(which(y)),]
  str_sub(a,explodemelocation[1],explodemelocation[2])<-"0"
  ### return
  a}
```

Snail Split:
if there's a number >=10, change it to [floor(x/2),ceiling(x/2)]
```{r}
snailsplit<-function(a){
  ### find a number greater than 10
  x<-str_locate(a,"[0-9][0-9]+")
  ### grab it
  y<-as.numeric(str_sub(a,x[1,1],x[1,2]))
  ### create the string that it splits to
  z<-str_c("[",floor(y/2),",",ceiling(y/2),"]")
  ### substitute that back into the string
  str_sub(a,x[1,1],x[1,2])<-z
  a}
```


Snail Magnitude
```{r}
snailmagnitude<-function(a){
  ## if there are any things of the form "[a,b]" (no nests,)
  while(str_detect(a,"\\[[0-9]+,[0-9]+\\]")){
    ### find where they are
    x<-str_locate(a,"\\[[0-9]+,[0-9]+\\]")
    ### pull that
    y<-str_sub(a,x[1,1],x[1,2])
    ### multiply a by 3 & b by 2
    z<-as.numeric(unlist(str_split(str_replace_all(y,"[\\[\\]]",""),",")))
    z<-3*z[1]+2*z[2]
    ### plug that into place where the nested value was
    str_sub(a,x[1,1],x[1,2])<-z}
  ## return the single number left.
  a}

```


So to find the result:
```{r}
part1<-input[1]
for(i in 2:length(input)){
  part1<-snailadd(part1,input[i])
  part1<-snailreduce(part1)}
part1
part1<-snailmagnitude(part1)
part1
```
## Part 2
Again, when all else fails - brute force it.

```{r}
part2df<-as.data.frame(matrix(ncol=2,nrow=0))
for(i in 1:length(input)){
  for(j in 1:length(input)){
    if(i!=j){part2df<-rbind(part2df,list(i,j))}}}
colnames(part2df)<-c("a","b")
```


```{r}
#948 is too low
part2df <- part2df %>% rowwise %>%
  mutate(mag=as.numeric(snailmagnitude(snailreduce(snailadd(input[a],input[b])))))
part2<-max(part2df$mag)
part2
```

