---
title: "Day 22 Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
library(knitr)
library(dplyr)
library(stringr)
library(tidyverse)
library(readr)
library(collections)
options(scipen = 999)
```

```{r}
input<-read_lines("Day22Sample2.txt")
```

## Part 1
Under "things that might take too much memory, but what the hell."
First, reparse the input so that it is of the form
c(on,xmin,xmax,ymin,ymax,zmin,zmax)


```{r}
rebootstep<-lapply(input,function(x){
  x<-str_squish(str_replace_all(x,"[x=|,y=|,z=|\\.\\.]"," "))
  x<-str_replace(x,"on","1")
  x<-str_replace(x,"off","0")
  x<-as.numeric(unlist(str_split(x," ")))
})
```


The idea here is to take (up to) six whacks at the original, paring down to the part that gets change (if anything)



slice it does the whacking in a single dimension, separating left, middle, and right.

```{r}
sliceit<-function(block,inst){
  if(block[1]>=inst[1]){l<-c(NA,NA)
  }else{l<-c(block[1],min(block[2],(inst[1]-1)))}
  if(block[1]>inst[2]||block[2]<inst[1]){m<-c(NA,NA)
  }else{m<-c(max(block[1],inst[1]),min(block[2],inst[2]))}
  if(block[2]<=inst[2]){h<-c(NA,NA)
  }else{h<-c(max(inst[2]+1,block[1]),block[2])}
c(l,m,h)}

```

chopblock does this in all dimensions.  It ends with up to seven updated (or not updated) new blocks.

```{r}
chopablock<-function(block,inst){
  outlist<-list()
  ### if this instruction is to turn something on and the block is already on, return the block.
  if(block[1]==inst[1]){outlist<-c(outlist,list(block))
  }else{
    #otherwise slice x
    x<-sliceit(block[2:3],inst[2:3])
    if(!is.na(x[1])){
      outlist<-c(outlist,list(c(block[1],x[1:2],block[4:7])))}
    if(!is.na(x[5])){
      outlist<-c(outlist,list(c(block[1],x[5:6],block[4:7])))}
    if(is.na(x[3])){block<-NA}else{block<-c(block[1],x[3:4],block[4:7])}    
    ### slice y
    if(any(!is.na(block))){
      y<-sliceit(block[4:5],inst[4:5])
      if(!is.na(y[1])){
        outlist<-c(outlist,list(c(block[1:3],y[1:2],block[6:7])))}
      if(!is.na(y[5])){
        outlist<-c(outlist,list(c(block[1:3],y[5:6],block[6:7])))}
      if(is.na(y[3])){block<-NA}else{block<-c(block[1:3],y[3:4],block[6:7])}
      ### if there is any block left, slice z
      if(any(!is.na(block))){
        z<-sliceit(block[6:7],inst[6:7])
        if(!is.na(z[1])){
          outlist<-c(outlist,list(c(block[1:5],z[1:2])))}
        if(!is.na(z[5])){
          outlist<-c(outlist,list(c(block[1:5],z[5:6])))}
        if(is.na(z[3])){block<-NA}else{block<-c(block[1:5],z[3:4])}
        ### assuming everything is sliced - add block
        if(any(!is.na(block))){
          block[1]<-1-block[1]
        outlist<-c(outlist,list(block))}}}}
outlist}
```

Now, we take our block (0,-50,50,-50,50,-50,50 ), and run it through instruction 1 to get a list of new blocks.  We run each of those blocks through inst 2 to get new blocks & so on & so on until we're out of instructions

```{r}
diceablock<-function(block, instructions){
  outblocks<-list(block)
  for(i in 1:length(instructions)){
    outblocks <- sapply(outblocks,function(x){
      ob<-chopablock(block=x,inst=instructions[[i]])
      ob})
    ### we need a list of vectors, so get a list of vectors
    if(class(outblocks[[1]])=="list"){outblocks<-unlist(outblocks,recursive = FALSE)}}
  outblocks}
  
```

```{r}
part1<-diceablock(c(0,-50,50,-50,50,-50,50),rebootstep)
```

Finally, we need to count how many are on:

```{r}
oncounter<-function(blocklist){
ons<-0
for(i in 1:length(blocklist)){
  block<-blocklist[[i]]
  if(block[1]==1){
    tot<-prod(abs(block[c(2,4,6)]-block[c(3,5,7)])+c(1,1,1))
    ons<-ons+tot}}
ons}

part1<-oncounter(part1)
part1
```
## Part 2

find the min & max of all dimensions to give the bounds of the block
```{r}
startingblock<-c(0,0,0,0,0,0)
for(i in 1:length(rebootstep)){
  x<-rebootstep[[i]]
  startingblock<-c(min(startingblock[1],x[2]),max(startingblock[2],x[3]),
                   min(startingblock[3],x[4]),max(startingblock[4],x[5]),
                   min(startingblock[5],x[6]),max(startingblock[6],x[7]))}
startingblock<-c(0,startingblock)
startingblock
```

Run with this bigger block.
```{r}

part2<-diceablock(startingblock,rebootstep)
part2<-oncounter(part2)
part2

```

