---
title: "Day 23 Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
library(knitr)
library(dplyr)
library(stringr)
library(tidyverse)
library(readr)
library(collections)
options(scipen = 999)
```


All of these functions take in a map and return a list of possible maps and the additional cost for one move past the  one that was brought in.

The hall functions move something in the hall to its appropriate room in position 1 (if position 1 is empty)

```{r}
hall1<-function(diag){
  hall<-diag[[1]]
  rooma<-diag[[2]]
  roomb<-diag[[3]]
  roomc<-diag[[4]]
  roomd<-diag[[5]]
  cost<-diag[[6]]
  ### move if position 1 in that room is 0.  
  switch(hall[1],
         "A"={
           if(rooma[1]=="."&&(all(rooma[-1]%in%c(".","A")))&&hall[2]=="."){cost<-cost+3
           hall[1]<-"."
           rooma[1]<-"A"}else{cost<-0}},
         "B"={
           if(roomb[1]=="."&&(all(roomb[-1]%in%c(".","B")))&&hall[2]=="."&&hall[3]=="."){cost<-cost+50
           hall[1]<-"."
           roomb[1]<-"B"}else{cost<-0}},
         "C"={
           if(roomc[1]=="."&&(all(roomc[-1]%in%c(".","C")))&&hall[2]=="."&&hall[3]=="."&&hall[4]=="."){
             cost<-cost+700
           hall[1]<-"."
           roomc[1]<-"C"}else{cost<-0}},
         "D"={
           if(roomd[1]=="."&&(all(roomd[-1]%in%c(".","D")))&&hall[2]=="."&&hall[3]=="."&&hall[4]=="."&&hall[5]=="."){cost<-cost+9000
           hall[1]<-"."
           roomd[1]<-"D"}else{cost<-0}},
         cat("something went wrong hall1\n"))
  ### return the updated diagram with the new cost
  return(list(c(hall),c(rooma),c(roomb),c(roomc),c(roomd),cost))}
####
hall2<-function(diag){
  hall<-diag[[1]]
  rooma<-diag[[2]]
  roomb<-diag[[3]]
  roomc<-diag[[4]]
  roomd<-diag[[5]]
  cost<-diag[[6]]
  switch(hall[2],
         "A"={
           if(rooma[1]=="."&&(all(rooma[-1]%in%c(".","A")))){cost<-cost+2
           hall[2]<-"."
           rooma[1]<-"A"}else{cost<-0}},
         "B"={
           if(roomb[1]=="."&&(all(roomb[-1]%in%c(".","B")))&&hall[3]=="."){cost<-cost+40
           hall[2]<-"."
           roomb[1]<-"B"}else{cost<-0}},
         "C"={
           if(roomc[1]=="."&&(all(roomc[-1]%in%c(".","C")))&&hall[3]=="."&&hall[4]=="."){cost<-cost+600
           hall[2]<-"."
           roomc[1]<-"C"}else{cost<-0}},
         "D"={
           if(roomd[1]=="."&&(all(roomd[-1]%in%c(".","D")))&&hall[3]=="."&&hall[4]=="."&&hall[5]=="."){cost<-cost+8000
           hall[2]<-"."
           roomd[1]<-"D"}else{cost<-0}},
         cat("something went wrong hall2\n"))
  return(list(c(hall),c(rooma),c(roomb),c(roomc),c(roomd),cost))}
####
hall3<-function(diag){
  hall<-diag[[1]]
  rooma<-diag[[2]]
  roomb<-diag[[3]]
  roomc<-diag[[4]]
  roomd<-diag[[5]]
  cost<-diag[[6]]
  switch(hall[3],
         "A"={
           if(rooma[1]=="."&&(all(rooma[-1]%in%c(".","A")))){cost<-cost+2
           hall[3]<-"."
           rooma[1]<-"A"}else{cost<-0}},
         "B"={
           if(roomb[1]=="."&&(all(roomb[-1]%in%c(".","B")))){cost<-cost+20
           hall[3]<-"."
           roomb[1]<-"B"}else{cost<-0}},
         "C"={
           if(roomc[1]=="."&&(all(roomc[-1]%in%c(".","C"))) && hall[4]=="."){cost<-cost+400
           hall[3]<-"."
           roomc[1]<-"C"}else{cost<-0}},
         "D"={
           if(roomd[1]=="."&&(all(roomd[-1]%in%c(".","D"))) &&hall[4]=="."&&hall[5]=="."){cost<-cost+6000
           hall[3]<-"."
           roomd[1]<-"D"}else{cost<-0}},
         cat("something went wrong hall3\n"))
  return(list(c(hall),c(rooma),c(roomb),c(roomc),c(roomd),cost))}
####
hall4<-function(diag){
  hall<-diag[[1]]
  rooma<-diag[[2]]
  roomb<-diag[[3]]
  roomc<-diag[[4]]
  roomd<-diag[[5]]
  cost<-diag[[6]]
  switch(hall[4],
         "A"={
           if(rooma[1]=="."&&(all(rooma[-1]%in%c(".","A")))&&hall[3]=="."){cost<-cost+4
           hall[4]<-"."
           rooma[1]<-"A"}else{cost<-0}},
         "B"={
           if(roomb[1]=="."&&(all(roomb[-1]%in%c(".","B")))){cost<-cost+20
           hall[4]<-"."
           roomb[1]<-"B"}else{cost<-0}},
         "C"={
           if(roomc[1]=="."&&(all(roomc[-1]%in%c(".","C")))){cost<-cost+200
           hall[4]<-"."
           roomc[1]<-"C"}else{cost<-0}},
         "D"={
           if(roomd[1]=="."&&(all(roomd[-1]%in%c(".","D")))&&hall[5]=="."){cost<-cost+4000
           hall[4]<-"."
           roomd[1]<-"D"}else{cost<-0}},
         cat("something went wrong hall4\n"))
  return(list(c(hall),c(rooma),c(roomb),c(roomc),c(roomd),cost))}
####
hall5<-function(diag){
  hall<-diag[[1]]
  rooma<-diag[[2]]
  roomb<-diag[[3]]
  roomc<-diag[[4]]
  roomd<-diag[[5]]
  cost<-diag[[6]]
  switch(hall[5],
         "A"={
           if(rooma[1]=="."&&(all(rooma[-1]%in%c(".","A")))&&hall[3]=="."&&hall[4]=="."){cost<-cost+6
           hall[5]<-"."
           rooma[1]<-"A"}else{cost<-0}},
         "B"={
           if(roomb[1]=="."&&(all(roomb[-1]%in%c(".","B")))&&hall[4]=="."){cost<-cost+40
           hall[5]<-"."
           roomb[1]<-"B"}else{cost<-0}},
         "C"={
           if(roomc[1]=="."&&(all(roomc[-1]%in%c(".","C")))){cost<-cost+200
           hall[5]<-"."
           roomc[1]<-"C"}else{cost<-0}},
         "D"={
           if(roomd[1]=="."&&(all(roomd[-1]%in%c(".","D")))){cost<-cost+2000
           hall[5]<-"."
           roomd[1]<-"D"}else{cost<-0}},
         cat("something went wrong hall5\n"))
  return(list(c(hall),c(rooma),c(roomb),c(roomc),c(roomd),cost))}
####
hall6<-function(diag){
  hall<-diag[[1]]
  rooma<-diag[[2]]
  roomb<-diag[[3]]
  roomc<-diag[[4]]
  roomd<-diag[[5]]
  cost<-diag[[6]]
  switch(hall[6],
         "A"={
           if(rooma[1]=="."&&(all(rooma[-1]%in%c(".","A")))&&hall[3]=="."&&hall[4]=="."&&hall[5]=="."){cost<-cost+8
           hall[6]<-"."
           rooma[1]<-"A"}else{cost<-0}},
         "B"={
           if(roomb[1]=="."&&(all(roomb[-1]%in%c(".","B")))&&hall[4]=="."&&hall[5]=="."){cost<-cost+60
           hall[6]<-"."
           roomb[1]<-"B"}else{cost<-0}},
         "C"={
           if(roomc[1]=="."&&(all(roomc[-1]%in%c(".","C")))&&hall[5]=="."){cost<-cost+400
           hall[6]<-"."
           roomc[1]<-"C"}else{cost<-0}},
         "D"={
           if(roomd[1]=="."&&(all(roomd[-1]%in%c(".","D")))){cost<-cost+2000
           hall[6]<-"."
           roomd[1]<-"D"}else{cost<-0}},
         cat("something went wrong hall6\n"))
  return(list(c(hall),c(rooma),c(roomb),c(roomc),c(roomd),cost))}
####
hall7<-function(diag){
 # cat(paste(unlist(diag)),"\n")
  hall<-diag[[1]]
  rooma<-diag[[2]]
  roomb<-diag[[3]]
  roomc<-diag[[4]]
  roomd<-diag[[5]]
  cost<-diag[[6]]
  switch(hall[7],
         "A"={
           if(rooma[1]=="."&&(all(rooma[-1]%in%c(".","A")))&&hall[3]=="."&&hall[4]=="."&&hall[5]=="."&&hall[6]=="."){cost<-cost+9
           hall[7]<-"."
           rooma[1]<-"A"}else{cost<-0}},
         "B"={
           if(roomb[1]=="."&&(all(roomb[-1]%in%c(".","B")))&&hall[4]=="."&&hall[5]=="."&&hall[6]=="."){cost<-cost+70
           hall[7]<-"."
           roomb[1]<-"B"}else{cost<-0}},
         "C"={
           if(roomc[1]=="."&&(all(roomc[-1]%in%c(".","C")))&&hall[5]=="."&&hall[6]=="."){cost<-cost+500
           hall[7]<-"."
           roomc[1]<-"C"}else{cost<-0}},
         "D"={
           if(roomd[1]=="."&&(all(roomd[-1]%in%c(".","D")))&&hall[6]=="."){cost<-cost+3000
           hall[7]<-"."
           roomd[1]<-"D"}else{cost<-0}},
         cat("something went wrong hall7\n"))
  return(list(c(hall),c(rooma),c(roomb),c(roomc),c(roomd),cost))}
```


The room function does flips the room to move correct amphipods deeper or incorrect amphipods to space 1.  It also puts amphipods in space 1 into all of the hallway positions

```{r}
### only go to the rooms if:
# need to move something deeper in e.g., rooma == c("A",".") move to c(".","A")
# need to move closer out, e.g., rooma==c(".","D") move to c("D",".")
# need to move something into the hall e.g. c("D",doesn't matter) move to c(".",still doesn't matter), hall = something new
rooma<-function(diag){
  #create the list of things to be returned
  outlist<-list()
  hall<-diag[[1]]
  rooma<-diag[[2]]
  roomb<-diag[[3]]
  roomc<-diag[[4]]
  roomd<-diag[[5]]
  oldcost<-diag[[6]]
  if(all(rooma==c("A","."))){
    outlist<-list(list(c(hall),rev(rooma),c(roomb),c(roomc),c(roomd),oldcost+1))
  }else if(rooma[1]=="."&&rooma[2]!="A"){
    switch(rooma[2],
           "B"=cost<-10,
           "C"=cost<-100,
           "D"=cost<-1000,
           cat("somethings wrong in rooma\n"))
    outlist<-list(list(c(hall),rev(rooma),c(roomb),c(roomc),c(roomd),oldcost+cost))
  }else{
    ### move to hallway
    ### to the left
    if(rooma[1]%in%c("A","B","C","D")){
      switch(rooma[1],
             "A"=cost<-1,
             "B"=cost<-10,
             "C"=cost<-100,
             "D"=cost<-1000,
             cat("somethings wrong in rooma\n"))
      if(hall[2]=="."){
        x<-list(c(hall[1],rooma[1],hall[3:7]),c(".",rooma[2]),c(roomb),c(roomc),c(roomd),oldcost+cost*3)
        outlist<-c(outlist,list(x))
        if(hall[1]=="."){
          x<-list(c(rooma[1],hall[2:7]),c(".",rooma[2]),c(roomb),c(roomc),c(roomd),oldcost+cost*2)
          outlist<-c(outlist,list(x))}}
      if(hall[3]=="."){
        x<-list(c(hall[1:2],rooma[1],hall[4:7]),c(".",rooma[2]),c(roomb),c(roomc),c(roomd),oldcost+cost*2)
        outlist<-c(outlist,list(x))
        if(hall[4]=="."){
          x<-list(c(hall[1:3],rooma[1],hall[5:7]),c(".",rooma[2]),c(roomb),c(roomc),c(roomd),oldcost+cost*4)
          outlist<-c(outlist,list(x))
          if(hall[5]=="."){
            x<-list(c(hall[1:4],rooma[1],hall[6:7]),c(".",rooma[2]),c(roomb),c(roomc),c(roomd),oldcost+cost*6)
            outlist<-c(outlist,list(x))
            if(hall[6]=="."){
              x<-list(c(hall[1:5],rooma[1],hall[7]),c(".",rooma[2]),c(roomb),c(roomc),c(roomd),oldcost+cost*8)
              outlist<-c(outlist,list(x))
              if(hall[7]=="."){
                x<-list(c(hall[1:6],rooma[1]),c(".",rooma[2]),c(roomb),c(roomc),c(roomd),oldcost+cost*9)
                outlist<-c(outlist,list(x))}}}}}}}
  outlist}

###

roomb<-function(diag){
  #create the list of things to be returned
  outlist<-list()
  hall<-diag[[1]]
  rooma<-diag[[2]]
  roomb<-diag[[3]]
  roomc<-diag[[4]]
  roomd<-diag[[5]]
  oldcost<-diag[[6]]
  if(all(roomb==c("B","."))){
    outlist<-list(list(c(hall),c(rooma),rev(roomb),c(roomc),c(roomd),oldcost+10))
  }else if(roomb[1]=="."&&roomb[2]!="B"){
    switch(roomb[2],
           "A"=cost<-1,
           "C"=cost<-100,
           "D"=cost<-1000,
           cat("somethings wrong in roomb\n"))
    outlist<-list(list(c(hall),c(rooma),rev(roomb),c(roomc),c(roomd),oldcost+cost))
  }else{
    ### move to hallway
    ### to the left
    if(roomb[1]%in%c("A","B","C","D")){
      switch(roomb[1],
             "A"=cost<-1,
             "B"=cost<-10,
             "C"=cost<-100,
             "D"=cost<-1000,
             cat("somethings wrong in roomb\n"))
      if(hall[3]=="."){
        x<-list(c(hall[1:2],roomb[1],hall[4:7]),c(rooma),c(".",roomb[2]),c(roomc),c(roomd),oldcost+cost*2)
        outlist<-c(outlist,list(x))
        if(hall[2]=="."){
          x<-list(c(hall[1],roomb[1],hall[3:7]),c(rooma),c(".",roomb[2]),c(roomc),c(roomd),oldcost+cost*4)
          outlist<-c(outlist,list(x))
          if(hall[1]=="."){
            x<-list(c(roomb[1],hall[2:7]),c(rooma),c(".",roomb[2]),c(roomc),c(roomd),oldcost+cost*5)
            outlist<-c(outlist,list(x))}}}
      if(hall[4]=="."){
        x<-list(c(hall[1:3],roomb[1],hall[5:7]),c(rooma),c(".",roomb[2]),c(roomc),c(roomd),oldcost+cost*2)
        outlist<-c(outlist,list(x))
        if(hall[5]=="."){
          x<-list(c(hall[1:4],roomb[1],hall[6:7]),c(rooma),c(".",roomb[2]),c(roomc),c(roomd),oldcost+cost*4)
          outlist<-c(outlist,list(x))
          if(hall[6]=="."){
            x<-list(c(hall[1:5],roomb[1],hall[7]),c(rooma),c(".",roomb[2]),c(roomc),c(roomd),oldcost+cost*6)
            outlist<-c(outlist,list(x))
            if(hall[7]=="."){
              x<-list(c(hall[1:6],roomb[1]),c(rooma),c(".",roomb[2]),c(roomc),c(roomd),oldcost+cost*7)
              outlist<-c(outlist,list(x))}}}}}}
  outlist}


#######

roomc<-function(diag){
  #create the list of things to be returned
  outlist<-list()
  hall<-diag[[1]]
  rooma<-diag[[2]]
  roomb<-diag[[3]]
  roomc<-diag[[4]]
  roomd<-diag[[5]]
  oldcost<-diag[[6]]
  if(all(roomc==c("C","."))){
    outlist<-list(list(c(hall),c(rooma),c(roomb),rev(roomc),c(roomd),oldcost+100))
  }else if(roomc[1]=="."&&roomc[2]!="C"){
    switch(roomc[2],
           "A"=cost<-1,
           "B"=cost<-10,
           "D"=cost<-1000,
           cat("somethings wrong in roomc\n"))
    outlist<-list(list(c(hall),c(rooma),c(roomb),rev(roomc),c(roomd),oldcost+cost))
  }else{
    ### move to hallway
    ### to the left
    if(roomc[1]%in%c("A","B","C","D")){
      switch(roomc[1],
             "A"=cost<-1,
             "B"=cost<-10,
             "C"=cost<-100,
             "D"=cost<-1000,
             cat("somethings wrong in roomc\n"))
      if(hall[4]=="."){
        x<-list(c(hall[1:3],roomc[1],hall[5:7]),c(rooma),c(roomb),c(".",roomc[2]),c(roomd),oldcost+cost*2)
        outlist<-c(outlist,list(x))
        if(hall[3]=="."){
          x<-list(c(hall[1:2],roomc[1],hall[4:7]),c(rooma),c(roomb),c(".",roomc[2]),c(roomd),oldcost+cost*4)
          outlist<-c(outlist,list(x))
          if(hall[2]=="."){
            x<-list(c(hall[1],roomc[1],hall[3:7]),c(rooma),c(roomb),c(".",roomc[2]),c(roomd),oldcost+cost*6)
            outlist<-c(outlist,list(x))
            if(hall[1]=="."){
              x<-list(c(roomc[1],hall[2:7]),c(rooma),c(roomb),c(".",roomc[2]),c(roomd),oldcost+cost*7)
              outlist<-c(outlist,list(x))}}}}
      if(hall[5]=="."){
        x<-list(c(hall[1:4],roomc[1],hall[6:7]),c(rooma),c(roomb),c(".",roomc[2]),c(roomd),oldcost+cost*2)
        outlist<-c(outlist,list(x))
        if(hall[6]=="."){
          x<-list(c(hall[1:5],roomc[1],hall[7]),c(rooma),c(roomb),c(".",roomc[2]),c(roomd),oldcost+cost*4)
          outlist<-c(outlist,list(x))
          if(hall[7]=="."){
            x<-list(c(hall[1:6],roomc[1]),c(rooma),c(roomb),c(".",roomc[2]),c(roomd),oldcost+cost*5)
            outlist<-c(outlist,list(x))}}}}}
  outlist}

##################

roomd<-function(diag){
  #create the list of things to be returned
  outlist<-list()
  hall<-diag[[1]]
  rooma<-diag[[2]]
  roomb<-diag[[3]]
  roomc<-diag[[4]]
  roomd<-diag[[5]]
  oldcost<-diag[[6]]
  if(all(roomd==c("D","."))){
    outlist<-list(list(c(hall),c(rooma),c(roomb),c(roomc),rev(roomd),oldcost+1000))
  }else if(roomd[1]=="."&&roomd[2]!="C"){
    switch(roomd[2],
           "A"=cost<-1,
           "B"=cost<-10,
           "C"=cost<-100,
           cat("somethings wrong in roomd p1\n"))
    outlist<-list(list(c(hall),c(rooma),c(roomb),c(roomc),rev(roomd),oldcost+cost))
  }else{
    ### move to hallway
    ### to the left
    if(roomd[1]%in%c("A","B","C","D")){
      switch(roomd[1],
             "A"=cost<-1,
             "B"=cost<-10,
             "C"=cost<-100,
             "D"=cost<-1000,
             cat("somethings wrong in roomd p2\n"))
      if(hall[5]=="."){
        x<-list(c(hall[1:4],roomd[1],hall[6:7]),c(rooma),c(roomb),c(roomc),c(".",roomd[2]),oldcost+cost*2)
        outlist<-c(outlist,list(x))
        if(hall[4]=="."){
          x<-list(c(hall[1:3],roomd[1],hall[5:7]),c(rooma),c(roomb),c(roomc),c(".",roomd[2]),oldcost+cost*4)
          outlist<-c(outlist,list(x))
          if(hall[3]=="."){
            x<-list(c(hall[1:2],roomd[1],hall[4:7]),c(rooma),c(roomb),c(roomc),c(".",roomd[2]),oldcost+cost*6)
            outlist<-c(outlist,list(x))
            if(hall[2]=="."){
              x<-list(c(hall[1],roomd[1],hall[3:7]),c(rooma),c(roomb),c(roomc),c(".",roomd[2]),oldcost+cost*8)
              outlist<-c(outlist,list(x))
              if(hall[1]=="."){
                x<-list(c(roomd[1],hall[2:7]),c(rooma),c(roomb),c(roomc),c(".",roomd[2]),oldcost+cost*9)
                outlist<-c(outlist,list(x))}}}}}
      if(hall[6]=="."){
        x<-list(c(hall[1:5],roomd[1],hall[7]),c(rooma),c(roomb),c(roomc),c(".",roomd[2]),oldcost+cost*2)
        outlist<-c(outlist,list(x))
        if(hall[7]=="."){
          x<-list(c(hall[1:6],roomd[1]),c(rooma),c(roomb),c(roomc),c(".",roomd[2]),oldcost+cost*3)
          outlist<-c(outlist,list(x))}}}}
  outlist}

```

The full function makes one move at a time. Either a hallway move or one or more room moves.

```{r}
amphipods<-function(a,b,c,d){
  pq<-priority_queue()
  beenthere<-dict()
  start<-list(rep(".",7),c(a),c(b),c(c),c(d),0)
  finalstate<-c(rep(".",7),c("A","A"),c("B","B"),c("C","C"),c("D","D"))
  pq$push(start,0)
  counter<-1
  while(pq$size()>0){
    now<-pq$pop()
    ### if done, done
    if(all(now[[1]]==".")&&all(now[[2]]=="A")){return(now)}
    ### if been there, next
    if(beenthere$has(now[1:5])){
#      counter<-counter+1
#      if(counter%%10000==0){cat(counter,now[[6]],pq$size(),beenthere$size(),"\n")}
      next}
    ### otherwise, here to the dictionary
    beenthere$set(now[1:5],now[[6]])
    #### Hallway moves
    if(now[[1]][1]!="."){
      h1<-hall1(now)
      if(h1[[6]]>0){pq$push(h1,-h1[[6]])}
    }
    if(now[[1]][2]!="."){
      h2<-hall2(now)
      if(h2[[6]]>0){pq$push(h2,-h2[[6]])}
    }
    if(now[[1]][3]!="."){
      h3<-hall3(now)
      if(h3[[6]]>0){pq$push(h3,-h3[[6]])}
    }
    if(now[[1]][4]!="."){
      h4<-hall4(now)
      if(h4[[6]]>0){pq$push(h4,-h4[[6]])}
    }
    if(now[[1]][5]!="."){
      h5<-hall5(now)
      if(h5[[6]]>0){pq$push(h5,-h5[[6]])}
    }
    if(now[[1]][6]!="."){
      h6<-hall6(now)
      if(h6[[6]]>0){pq$push(h6,-h6[[6]])}
    }
    if(now[[1]][7]!="."){
      h7<-hall7(now)
      if(h7[[6]]>0){pq$push(h7,-h7[[6]])}
    }
    ### room moves
    if(any(now[[2]]!=".")&& !all(now[[2]]==c(".","A"))){
      ra<-rooma(now)
      if(length(ra)>0){
        for(i in 1:length(ra)){
          x<-ra[[i]]
          pq$push(x,-x[[6]])}}}
    if(any(now[[3]]!=".")&& !all(now[[3]]==c(".","B"))){
      rb<-roomb(now)
      if(length(rb)>0){
        for(i in 1:length(rb)){
          x<-rb[[i]]
          pq$push(x,-x[[6]])
          }}}
    if(any(now[[4]]!=".")&& !all(now[[4]]==c(".","C"))){
      rc<-roomc(now)
      if(length(rc)>0){
        for(i in 1:length(rc)){
          x<-rc[[i]]
          pq$push(x,-x[[6]])}}}
    if(any(now[[5]]!=".")&& !all(now[[5]]==c(".","D"))){
      rd<-roomd(now)
      if(length(rd)>0){
        for(i in 1:length(rd)){
          x<-rd[[i]]
          #cat("d",paste(x))    
          pq$push(x,-x[[6]])
          }}}
    
    
    
    
  }
  return(beenthere)}
```


Run the function with input
```{r}
### Puzzle input is done manually

a<-c("B","A")
b<-c("C","D")
c<-c("B","C")
d<-c("D","A")

p1<-amphipods(a,b,c,d)
part1<-p1[[6]]
part1
```


## Part 2

For the halls, nothing changes.  They're already set up to to be called & to work no matter how long the rooms are.

The rooms need to be split into three functions for my sanity.

Nothing really changes in the room to hall - they are split apart from the other room functions

```{r}
### if there is something in roomX[1]&&if there is space in the hall&&if there is something in the room that belongs in a different room:
roomouta<-function(diag){
  #create the list of things to be returned
  outlist<-list()
  hall<-diag[[1]]
  rooma<-diag[[2]]
  roomb<-diag[[3]]
  roomc<-diag[[4]]
  roomd<-diag[[5]]
  oldcost<-diag[[6]]
  ### move to hallway
  ### to the left
 # cat(paste(diag,"\n"))
  switch(rooma[1],
         "A"=cost<-1,
         "B"=cost<-10,
         "C"=cost<-100,
         "D"=cost<-1000,
         cat("somethings wrong in rooma\n"))
  if(hall[2]=="."){
    x<-list(c(hall[1],rooma[1],hall[3:7]),c(".",rooma[-1]),c(roomb),c(roomc),c(roomd),oldcost+cost*3)
    outlist<-c(outlist,list(x))
    if(hall[1]=="."){
      x<-list(c(rooma[1],hall[2:7]),c(".",rooma[-1]),c(roomb),c(roomc),c(roomd),oldcost+cost*2)
      outlist<-c(outlist,list(x))}}
  if(hall[3]=="."){
    x<-list(c(hall[1:2],rooma[1],hall[4:7]),c(".",rooma[-1]),c(roomb),c(roomc),c(roomd),oldcost+cost*2)
    outlist<-c(outlist,list(x))
    if(hall[4]=="."){
      x<-list(c(hall[1:3],rooma[1],hall[5:7]),c(".",rooma[-1]),c(roomb),c(roomc),c(roomd),oldcost+cost*4)
      outlist<-c(outlist,list(x))
      if(hall[5]=="."){
        x<-list(c(hall[1:4],rooma[1],hall[6:7]),c(".",rooma[-1]),c(roomb),c(roomc),c(roomd),oldcost+cost*6)
        outlist<-c(outlist,list(x))
        if(hall[6]=="."){
          x<-list(c(hall[1:5],rooma[1],hall[7]),c(".",rooma[-1]),c(roomb),c(roomc),c(roomd),oldcost+cost*8)
          outlist<-c(outlist,list(x))
          if(hall[7]=="."){
            x<-list(c(hall[1:6],rooma[1]),c(".",rooma[-1]),c(roomb),c(roomc),c(roomd),oldcost+cost*9)
            outlist<-c(outlist,list(x))}}}}}
outlist}

###

roomoutb<-function(diag){
  #create the list of things to be returned
  outlist<-list()
  hall<-diag[[1]]
  rooma<-diag[[2]]
  roomb<-diag[[3]]
  roomc<-diag[[4]]
  roomd<-diag[[5]]
  oldcost<-diag[[6]]
  switch(roomb[1],
         "A"=cost<-1,
         "B"=cost<-10,
         "C"=cost<-100,
         "D"=cost<-1000,
         cat("somethings wrong in roomb\n"))
  if(hall[3]=="."){
    x<-list(c(hall[1:2],roomb[1],hall[4:7]),c(rooma),c(".",roomb[-1]),c(roomc),c(roomd),oldcost+cost*2)
    outlist<-c(outlist,list(x))
    if(hall[2]=="."){
      x<-list(c(hall[1],roomb[1],hall[3:7]),c(rooma),c(".",roomb[-1]),c(roomc),c(roomd),oldcost+cost*4)
      outlist<-c(outlist,list(x))
      if(hall[1]=="."){
        x<-list(c(roomb[1],hall[2:7]),c(rooma),c(".",roomb[-1]),c(roomc),c(roomd),oldcost+cost*5)
        outlist<-c(outlist,list(x))}}}
  if(hall[4]=="."){
    x<-list(c(hall[1:3],roomb[1],hall[5:7]),c(rooma),c(".",roomb[-1]),c(roomc),c(roomd),oldcost+cost*2)
    outlist<-c(outlist,list(x))
    if(hall[5]=="."){
      x<-list(c(hall[1:4],roomb[1],hall[6:7]),c(rooma),c(".",roomb[-1]),c(roomc),c(roomd),oldcost+cost*4)
      outlist<-c(outlist,list(x))
      if(hall[6]=="."){
        x<-list(c(hall[1:5],roomb[1],hall[7]),c(rooma),c(".",roomb[-1]),c(roomc),c(roomd),oldcost+cost*6)
        outlist<-c(outlist,list(x))
        if(hall[7]=="."){
          x<-list(c(hall[1:6],roomb[1]),c(rooma),c(".",roomb[-1]),c(roomc),c(roomd),oldcost+cost*7)
          outlist<-c(outlist,list(x))}}}}
outlist}


#######

roomoutc<-function(diag){
  #create the list of things to be returned
  outlist<-list()
  hall<-diag[[1]]
  rooma<-diag[[2]]
  roomb<-diag[[3]]
  roomc<-diag[[4]]
  roomd<-diag[[5]]
  oldcost<-diag[[6]]
    ### move to hallway
    ### to the left
      switch(roomc[1],
             "A"=cost<-1,
             "B"=cost<-10,
             "C"=cost<-100,
             "D"=cost<-1000,
             cat("somethings wrong in roomc\n"))
      if(hall[4]=="."){
        x<-list(c(hall[1:3],roomc[1],hall[5:7]),c(rooma),c(roomb),c(".",roomc[-1]),c(roomd),oldcost+cost*2)
        outlist<-c(outlist,list(x))
        if(hall[3]=="."){
          x<-list(c(hall[1:2],roomc[1],hall[4:7]),c(rooma),c(roomb),c(".",roomc[-1]),c(roomd),oldcost+cost*4)
          outlist<-c(outlist,list(x))
          if(hall[2]=="."){
            x<-list(c(hall[1],roomc[1],hall[3:7]),c(rooma),c(roomb),c(".",roomc[-1]),c(roomd),oldcost+cost*6)
            outlist<-c(outlist,list(x))
            if(hall[1]=="."){
              x<-list(c(roomc[1],hall[2:7]),c(rooma),c(roomb),c(".",roomc[-1]),c(roomd),oldcost+cost*7)
              outlist<-c(outlist,list(x))}}}}
      if(hall[5]=="."){
        x<-list(c(hall[1:4],roomc[1],hall[6:7]),c(rooma),c(roomb),c(".",roomc[-1]),c(roomd),oldcost+cost*2)
        outlist<-c(outlist,list(x))
        if(hall[6]=="."){
          x<-list(c(hall[1:5],roomc[1],hall[7]),c(rooma),c(roomb),c(".",roomc[-1]),c(roomd),oldcost+cost*4)
          outlist<-c(outlist,list(x))
          if(hall[7]=="."){
            x<-list(c(hall[1:6],roomc[1]),c(rooma),c(roomb),c(".",roomc[-1]),c(roomd),oldcost+cost*5)
            outlist<-c(outlist,list(x))}}}
  outlist}

##################

roomoutd<-function(diag){
  #create the list of things to be returned
  outlist<-list()
  hall<-diag[[1]]
  rooma<-diag[[2]]
  roomb<-diag[[3]]
  roomc<-diag[[4]]
  roomd<-diag[[5]]
  oldcost<-diag[[6]]
    ### move to hallway
    ### to the left
      switch(roomd[1],
             "A"=cost<-1,
             "B"=cost<-10,
             "C"=cost<-100,
             "D"=cost<-1000,
             cat("somethings wrong in roomd p2\n"))
      if(hall[5]=="."){
        x<-list(c(hall[1:4],roomd[1],hall[6:7]),c(rooma),c(roomb),c(roomc),c(".",roomd[-1]),oldcost+cost*2)
        outlist<-c(outlist,list(x))
        if(hall[4]=="."){
          x<-list(c(hall[1:3],roomd[1],hall[5:7]),c(rooma),c(roomb),c(roomc),c(".",roomd[-1]),oldcost+cost*4)
          outlist<-c(outlist,list(x))
          if(hall[3]=="."){
            x<-list(c(hall[1:2],roomd[1],hall[4:7]),c(rooma),c(roomb),c(roomc),c(".",roomd[-1]),oldcost+cost*6)
            outlist<-c(outlist,list(x))
            if(hall[2]=="."){
              x<-list(c(hall[1],roomd[1],hall[3:7]),c(rooma),c(roomb),c(roomc),c(".",roomd[-1]),oldcost+cost*8)
              outlist<-c(outlist,list(x))
              if(hall[1]=="."){
                x<-list(c(roomd[1],hall[2:7]),c(rooma),c(roomb),c(roomc),c(".",roomd[-1]),oldcost+cost*9)
                outlist<-c(outlist,list(x))}}}}}
      if(hall[6]=="."){
        x<-list(c(hall[1:5],roomd[1],hall[7]),c(rooma),c(roomb),c(roomc),c(".",roomd[-1]),oldcost+cost*2)
        outlist<-c(outlist,list(x))
        if(hall[7]=="."){
          x<-list(c(hall[1:6],roomd[1]),c(rooma),c(roomb),c(roomc),c(".",roomd[-1]),oldcost+cost*3)
          outlist<-c(outlist,list(x))}}
  outlist}
```

The room forward should be called if there's nothing in space 1 and an incorrect amphipod somewhere further back.  It moves the closest to the front all the way to the front.  There they can be picked up by the room out function

```{r}
roomaforward<-function(diag){
  #create the list of things to be returned
  outlist<-diag
  hall<-diag[[1]]
  rooma<-diag[[2]]
  roomb<-diag[[3]]
  roomc<-diag[[4]]
  roomd<-diag[[5]]
  oldcost<-diag[[6]]
  ### find the first non empty space
  x<-min(which(rooma!="."))
  switch(rooma[x],
         "A"=cost<-1,
         "B"=cost<-10,
         "C"=cost<-100,
         "D"=cost<-1000,
         cat("somethings wrong in roomd forward\n"))
  ### put it in space 1 and make its space empty
  rooma[1]<-rooma[x]
  rooma[x]<-"."
  ### update the cost based on what was being moved.
  outlist<-list(c(hall),c(rooma),c(roomb),c(roomc),c(roomd),oldcost+((x-1)*cost))
  outlist}
####
roombforward<-function(diag){
  #create the list of things to be returned
  outlist<-diag
  hall<-diag[[1]]
  rooma<-diag[[2]]
  roomb<-diag[[3]]
  roomc<-diag[[4]]
  roomd<-diag[[5]]
  oldcost<-diag[[6]]
  x<-min(which(roomb!="."))
  switch(roomb[x],
         "A"=cost<-1,
         "B"=cost<-10,
         "C"=cost<-100,
         "D"=cost<-1000,
         cat("somethings wrong in roomd forward\n"))
  roomb[1]<-roomb[x]
  roomb[x]<-"."
  outlist<-list(c(hall),c(rooma),c(roomb),c(roomc),c(roomd),oldcost+((x-1)*cost))
  outlist}
####
roomcforward<-function(diag){
  #create the list of things to be returned
  outlist<-diag
  hall<-diag[[1]]
  rooma<-diag[[2]]
  roomb<-diag[[3]]
  roomc<-diag[[4]]
  roomd<-diag[[5]]
  oldcost<-diag[[6]]
  x<-min(which(roomc!="."))
  switch(roomc[x],
         "A"=cost<-1,
         "B"=cost<-10,
         "C"=cost<-100,
         "D"=cost<-1000,
         cat("somethings wrong in roomd forward\n"))
  roomc[1]<-roomc[x]
  roomc[x]<-"."
  outlist<-list(c(hall),c(rooma),c(roomb),c(roomc),c(roomd),oldcost+((x-1)*cost))
  outlist}
####
roomdforward<-function(diag){
  #create the list of things to be returned
  outlist<-diag
  hall<-diag[[1]]
  rooma<-diag[[2]]
  roomb<-diag[[3]]
  roomc<-diag[[4]]
  roomd<-diag[[5]]
  oldcost<-diag[[6]]
  x<-min(which(roomd!="."))
  switch(roomd[x],
         "A"=cost<-1,
         "B"=cost<-10,
         "C"=cost<-100,
         "D"=cost<-1000,
         cat("somethings wrong in roomd forward\n"))
  roomd[1]<-roomd[x]
  roomd[x]<-"."
  outlist<-list(c(hall),c(rooma),c(roomb),c(roomc),c(roomd),oldcost+((x-1)*cost))
  outlist}
```
Room back moves correct amphipods from position 1 (where they were dropped off by the hallway function) to the first space in the back of the room.

```{r}
roomaback<-function(diag){
  #create the list of things to be returned
  outlist<-diag
  hall<-diag[[1]]
  rooma<-diag[[2]]
  roomb<-diag[[3]]
  roomc<-diag[[4]]
  roomd<-diag[[5]]
  oldcost<-diag[[6]]
  ### find the first empty space
  x<-max(which(rooma=="."))
  ### swap it with the amphipod
  rooma[1]<-"."
  rooma[x]<-"A"
  ### update the cost
  outlist<-list(c(hall),c(rooma),c(roomb),c(roomc),c(roomd),oldcost+(x-1))
  outlist}
####
roombback<-function(diag){
  #create the list of things to be returned
  outlist<-diag
  hall<-diag[[1]]
  rooma<-diag[[2]]
  roomb<-diag[[3]]
  roomc<-diag[[4]]
  roomd<-diag[[5]]
  oldcost<-diag[[6]]
  x<-max(which(roomb=="."))
  roomb[1]<-"."
  roomb[x]<-"B"
  outlist<-list(c(hall),c(rooma),c(roomb),c(roomc),c(roomd),oldcost+10*(x-1))
  outlist}
####
roomcback<-function(diag){
  #create the list of things to be returned
  outlist<-diag
  hall<-diag[[1]]
  rooma<-diag[[2]]
  roomb<-diag[[3]]
  roomc<-diag[[4]]
  roomd<-diag[[5]]
  oldcost<-diag[[6]]
  x<-max(which(roomc=="."))
  roomc[1]<-"."
  roomc[x]<-"C"
  outlist<-list(c(hall),c(rooma),c(roomb),c(roomc),c(roomd),oldcost+100*(x-1))
  outlist}
####
roomdback<-function(diag){
  #create the list of things to be returned
  outlist<-diag
  hall<-diag[[1]]
  rooma<-diag[[2]]
  roomb<-diag[[3]]
  roomc<-diag[[4]]
  roomd<-diag[[5]]
  oldcost<-diag[[6]]
  x<-max(which(roomd=="."))
  roomd[1]<-"."
  roomd[x]<-"D"
  outlist<-list(c(hall),c(rooma),c(roomb),c(roomc),c(roomd),oldcost+1000*(x-1))
  outlist}
```


The whole function works the same way as it did before.

```{r}
amphipodslong<-function(a,b,c,d){
  pq<-priority_queue()
  beenthere<-dict()
  start<- list(rep(".",7),c(a),c(b),c(c),c(d),0)
  finalstate<-c(rep(".",7),c("A","A"),c("B","B"),c("C","C"),c("D","D"))
  pq$push(start,0)
  counter<-1
  while(pq$size()>0){
    now<-pq$pop()
    ### if done, done return the whole diagram for checking
    if(all(now[[1]]==".")&&all(now[[2]]=="A")&&all(now[[3]]=="B")){return(now)}
    ### if been there, next
    if(beenthere$has(now[1:5])){
      next}
    ### otherwise, add this to the dictionary
    beenthere$set(now[1:5],now[[6]])
    #### add in the hallway moves
    if(now[[1]][1]!="."){
      h1<-hall1(now)
      pq$push(h1,-h1[[6]])}
    if(now[[1]][2]!="."){
      h2<-hall2(now)
      pq$push(h2,-h2[[6]])}
    if(now[[1]][3]!="."){
      h3<-hall3(now)
      pq$push(h3,-h3[[6]])}
    if(now[[1]][4]!="."){
      h4<-hall4(now)
      pq$push(h4,-h4[[6]])}
    if(now[[1]][5]!="."){
      h5<-hall5(now)
      pq$push(h5,-h5[[6]])}
    if(now[[1]][6]!="."){
      h6<-hall6(now)
      pq$push(h6,-h6[[6]])}
    if(now[[1]][7]!="."){
      h7<-hall7(now)
      pq$push(h7,-h7[[6]])}
    ### First, Room outs
    ### add in the room outs for everything in position 1 where there is a lost amphipod
    if(any(now[[1]]==".")){
      if(now[[2]][1]!="."&&!all(now[[2]]%in%c(".","A"))){
        ra<-roomouta(now)
        if(length(ra)>0){
          for(i in 1:length(ra)){
            x<-ra[[i]]
            pq$push(x,-x[[6]])}}}
      if(now[[3]][1]!="."&&!all(now[[3]]%in%c(".","B"))){
        rb<-roomoutb(now)
        if(length(rb)>0){
          for(i in 1:length(rb)){
            x<-rb[[i]]
            pq$push(x,-x[[6]])}}}
      if(now[[4]][1]!="."&&!all(now[[4]]%in%c(".","C"))){
        rc<-roomoutc(now)
        if(length(rc)>0){
          for(i in 1:length(rc)){
            x<-rc[[i]]
            pq$push(x,-x[[6]])}}}
      if(now[[5]][1]!="."&&!all(now[[5]]%in%c(".","D"))){
        rd<-roomoutd(now)
        if(length(rd)>0){
          for(i in 1:length(rd)){
            x<-rd[[i]]
            pq$push(x,-x[[6]])}}}}
    ### shuffle amphipods forward to space 1
    if(now[[2]][1]=="."&&any(c("B","C","D")%in%now[[2]])){
      raf<-roomaforward(now)
      pq$push(raf,-raf[[6]])}
    if(now[[3]][1]=="."&&any(c("A","C","D")%in%now[[3]])){
      rbf<-roombforward(now)
      pq$push(rbf,-rbf[[6]])}
    if(now[[4]][1]=="."&&any(c("A","B","D")%in%now[[4]])){
      rcf<-roomcforward(now)
      pq$push(rcf,-rcf[[6]])}
    if(now[[5]][1]=="."&&any(c("A","B","C")%in%now[[5]])){
      rdf<-roomdforward(now)
      pq$push(rdf,-rdf[[6]])}
    ### shuffle found amphipods back to the back of the room.
    if(now[[2]][1]=="A"&&all(c(".","A")%in%now[[2]])&&all(now[[2]]%in%c(".","A"))){
      rab<-roomaback(now)
      pq$push(rab,-rab[[6]])}
    if(now[[3]][1]=="B"&&all(c(".","B")%in%now[[3]])&&all(now[[3]]%in%c(".","B"))){
      rbb<-roombback(now)
      pq$push(rbb,-rbb[[6]])}
    if(now[[4]][1]=="C"&&all(c(".","C")%in%now[[4]])&&all(now[[4]]%in%c(".","C"))){
      rcb<-roomcback(now)
      pq$push(rcb,-rcb[[6]])}
    if(now[[5]][1]=="D"&&all(c(".","D")%in%now[[5]])&&all(now[[5]]%in%c(".","D"))){
      rdb<-roomdback(now)
      pq$push(rdb,-rdb[[6]])}
    }
  return(list(now,beenthere))}
```


Now run - 
```{r}
### First show it works on the previous input (rooms size 2):

a<-c("B","A")
b<-c("C","D")
c<-c("B","C")
d<-c("D","A")
p1<-amphipodslong(a,b,c,d)
part1long<-p1[[6]]
part1long

### then that it works on the long input (room size 4)

a<-c("B","D","D","A")
b<-c("C","C","B","D")
c<-c("B","B","A","C")
d<-c("D","A","C","A")
p2<-amphipodslong(a,b,c,d)
part2<-p2[[6]]
part2
```