---
title: "Day 16 Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
library(knitr)
library(dplyr)
library(stringr)
library(tidyverse)
library(readr)
library(collections)
options(scipen = 999)
```

```{r}
input<-read_lines("Day16Sample.txt")
```

## Part 1

First, a binary to decimal converter (will be needed later)
```{r}
b2d<-function(s){
  x<-0
  for(i in 1:nchar(s)){
    x<-2*x+as.numeric(str_sub(s,i,i))}
  x}
```



First, convert the input to binary

```{r}
expanded<-sapply(input,function(x){
  z<-c("")
  for(i in 1:nchar(x)){
    y<-switch(str_sub(x,i,i),
              "0"="0000",
              "1"="0001",
              "2"="0010",
              "3"="0011",
              "4"="0100",
              "5"="0101",
              "6"="0110",
              "7"="0111",
              "8"="1000",
              "9"="1001",
              "A"="1010",
              "B"="1011",
              "C"="1100",
              "D"="1101",
              "E"="1110",
              "F"="1111",
              cat("something went wrong"))
    z<-str_c(z,y)}
  z})

```

Then parse this thing

First a function for when type = 4 (100 binary)

```{r}
type4<-function(st){
  litnum<-""
  keepgoing<-TRUE
  while(keepgoing){
    ### check the first character
    f<-str_sub(st,1,1)
    ### add next 4 characters to the number
    litnum<-str_c(litnum,str_sub(st,2,5))
    ### remove those from st
    st<-str_sub(st,6)
    ### if the first character was a 0, stop. Otherwise repeat
    if(f==0){keepgoing<-FALSE}}
  litnum<-b2d(litnum)
  ### return the number & the remainder of the string
  return(list(litnum,st))}
```

A function for when when the length type id=15

```{r}
operator0<-function(st){
  ### find how many characters are in the packet
  l<-b2d(str_sub(st,1,15))
  ### update the string
  st<-str_sub(st,16)
  #### get the packet's characters
  sp<-str_sub(st,1,l)
  ### update the string
  st<-str_sub(st,l+1)
  ### return the packet - run to the end of the packet.
  return(list(parseit(sp),st))}
```

A function for when the length type id=11

```{r}
operator1<-function(st){
  ### find out how many repeats
  l<-b2d(str_sub(st,1,11))
  ### update the string
  st<-str_sub(st,12)
  ### run *exactly* that many packets
  subpackets<-list()
  for(i in 1:l){
    x<-singlesubp(st)
    ### add that single subpacket to the list
    subpackets<-c(subpackets,list(x[[1]]))
    ### update st
    st<-x[[2]]}
  ### return the list of subpackets
  return(list(subpackets,st))}
  
  
```

Then a function to run exactly one packet
```{r}
singlesubp<-function(st){
  ### get version number
  v<-str_sub(st,1,3)
  v<-b2d(v)
  v<-str_c("v=",v)
  ### get type
  t<-str_sub(st,4,6)
  st<-str_sub(st,7)
  ### if type is 100 - run the type function & get packets
  if(t=="100"){
    x<-type4(st)
    sbp<-list(v,t,x[[1]])
    st<-x[[2]]
  }else{
    ### get the indicator
    i<-str_sub(st,1,1)
    st<-str_sub(st,2)
    if(i==0){
    ### if the indicator is 0, run operator 0 function
      x<-operator0(st)
      sbp<-list(v,t,x[[1]])
      st<-x[[2]]
    }else{
      ### if the indicator is 1, run the operator1 function
      x<-operator1(st)
      sbp<-list(v,t,x[[1]])
      st<-x[[2]]
      }
    }
  list(sbp,st)}

```


Then a function that keeps running them over and over and over.

```{r}
parseit<-function(st){
  ### if it's an empty string, then I do nothing.
  if(nchar(st)==0){}
  ### if there's no 1's, do nothing
  if(str_detect(st,"1",negate=TRUE)){}
  ### while there's string left, I need to create subpackets, unless I don't.
  subpackets<-list()
  ### create a packet
  while(nchar(st)>0&str_detect(st,"1")){
    x<-singlesubp(st)
    subpackets<-c(subpackets,list(x[[1]]))
    st<-x[[2]]}
subpackets}
```


For a single answer
```{r}
# p1<-parseit(expanded)
### unlist the list - just to find the v's
# p1<-unlist(p1)
### anything that has "v=" find the number associated with it.  Those are my versions
# part1<-sum(sapply(p1[which(str_detect(p1,"v="))],function(x){as.numeric(str_remove(x,"v="))}))
# part1
```

```{r}
part1<-sapply(expanded,function(x){
  y<-parseit(x)
  y<-unlist(y)
  y<-sapply(y[which(str_detect(y,"v="))],function(z){as.numeric(str_remove(z,"v="))})
  y<-sum(y)
  y})
part1
```




## Part 2

Start by writing a bin2dec style function, because one of my binary numbers is too long for the one that comes in the package.

```{r}
b2d<-function(s){
  x<-0
  for(i in 1:nchar(s)){
    x<-2*x+as.numeric(str_sub(s,i,i))}
  x}
```



Then *slightly* changing the functions.   This time, instead of just doing a list, they'll return an actual number based on being a literal or an operation



```{r}
type4_2<-function(st){
  litnum<-""
  keepgoing<-TRUE
  while(keepgoing){
    ### check the first character
    f<-str_sub(st,1,1)
    ### add next 4 characters to the number
    litnum<-str_c(litnum,str_sub(st,2,5))
    ### remove those from st
    st<-str_sub(st,6)
    ### if the first character was a 0, stop. Otherwise repeat
    if(f==0){keepgoing<-FALSE}}
  qqq<-litnum
  litnum<-b2d(litnum)
  if(is.na(litnum)){cat("!!!!!! H!H!H!HH!H",qqq,"\n")}
  ### return the number & the remainder of the string
  return(list(as.numeric(litnum),st))}
```

A function for when when the length type id=15

```{r}
operator0_2<-function(st){
  ### find how many characters are in the packet
  l<-b2d(str_sub(st,1,15))
  ### update the string
  st<-str_sub(st,16)
  #### get the packet's characters
  sp<-str_sub(st,1,l)
  ### update the string
  st<-str_sub(st,l+1)
  ### return the packet - run to the end of the packet.
  return(list(parseit_2(sp),st))}
```

A function for when the length type id=11

```{r}
operator1_2<-function(st){
  ### find out how many repeats
  l<-b2d(str_sub(st,1,11))
  ### update the string
  st<-str_sub(st,12)
  ### run *exactly* that many packets
  subpackets<-list()
  for(i in 1:l){
    x<-singlesubp_2(st)
    ### add that single subpacket to the list
    subpackets<-c(subpackets,list(x[[1]]))
    ### update st
    st<-x[[2]]}
  ### return the list of subpackets
  return(list(subpackets,st))}
  
  
```

Then a function to run exactly one packet
```{r}
singlesubp_2<-function(st){
  ### get version number
  v<-str_sub(st,1,3)
  v<-b2d(v)
  v<-str_c("v=",v)
  ### get type
  t<-str_sub(st,4,6)
  st<-str_sub(st,7)
  ### if type is 100 - run the type function & get packets.
  switch(t,
         "000"=oper<-"sum",
         "001"=oper<-"prod",
         "010"=oper<-"min",
         "011"=oper<-"max",
         "100"=oper<-"nada",
         "101"=oper<-">",
         "110"=oper<-"<",
         "111"=oper<-"==",
         cat("something went wrong\n"))
  if(t=="100"){
    x<-type4_2(st)
    ### if t=100 return the number
    sbp<-x[[1]]
    st<-x[[2]]
  }else{
    ### get the indicator
    i<-str_sub(st,1,1)
    st<-str_sub(st,2)
    if(i==0){
    ### if the indicator is 0, run operator 0 function
      x<-operator0_2(st)
      ### spb - run the operation
      sbp<-as.numeric(do.call(oper,x[[1]]))
      st<-x[[2]]
    }else{
      ### if the indicator is 1, run the operator1 function
      x<-operator1_2(st)
      ### spb - run the operation the "as numeric" makes TRUE=1,FALSE=0
      sbp<-as.numeric(do.call(oper,x[[1]]))
      st<-x[[2]]
      }}
  list(sbp,st)}

```


Then a function that keeps running them over and over and over.

```{r}
parseit_2<-function(st){
  ### if it's an empty string, then I do nothing.
  if(nchar(st)==0){}
  ### if there's no 1's, do nothing
  if(str_detect(st,"1",negate=TRUE)){}
  ### while there's string left, I need to create subpackets, unless I don't.
  subpackets<-list()
  ### create a packet
  while(nchar(st)>0&str_detect(st,"1")){
    x<-singlesubp_2(st)
    subpackets<-c(subpackets,list(x[[1]]))
    st<-x[[2]]}
subpackets}
```



```{r}
part2<-sapply(expanded,function(x){parseit_2(x)})
part2
```

