---
title: "Day 9 Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
library(igraph)
library(ggplot2)
library(rgl)
library(ggraph)
library(knitr)
library(dplyr)
library(stringr)
library(tidyverse)
library(readr)
library(collections)
library(reshape2)
options(scipen = 999)
```

```{r}
input<-read_lines("Day9Sample.txt")
### create the heightmap with a rim of 10 around the edges.
heightmap<-matrix(0,ncol=nchar(input[1])+2,nrow=length(input)+2)
for(i in 1:length(input)){heightmap[i+1,]<-c(10,as.numeric(unlist(str_split(input[i],""))),10)}
heightmap[1,]<-rep(10,ncol(heightmap))
heightmap[nrow(heightmap),]<-rep(10,ncol(heightmap))
```


## Part 1

Check every place in the heightmap to see if its neighbors are bigger.  If all 4 are, then add it (and its height) to the answer

```{r}
part1<-0
for(i in 2:(ncol(heightmap)-1)){
  for(j in 2:(nrow(heightmap)-1)){
    lp<-heightmap[j,i]
    if(lp<heightmap[j+1,i]&&
       lp<heightmap[j-1,i]&&
       lp<heightmap[j,i+1]&&
       lp<heightmap[j,i-1]){part1<-part1+1+lp}}}
part1


```

## Part 2

Find all of the flood paths


The flood fill function:

```{r}

### bring in x,y, the dictionary, and the height map
basins<-function(x,y,bt,hm){
  ### everything in this fill will be labeled with the starting point
  s<-paste(x,y)
  stk<-stack()
  ### push
  stk$push(c(x,y))
  while(stk$size()>0){
    curr<-stk$pop()
    ### remove anything that's already been filled, that's off the edge, or that =9 (if it hasn't already been found)
    if(bt$has(curr)){next}
    if(curr[1]<1||curr[2]<1||curr[1]>ncol(hm)||curr[2]>nrow(hm)){next}
    if(hm[curr[2],curr[1]]>=9){next}
    bt$set(curr,s)
    ### find the thing up, down, left, or right
    u<-c(curr[1],curr[2])+c(0,-1)
    d<-c(curr[1],curr[2])+c(0,1)
    r<-c(curr[1],curr[2])+c(1,0)
    l<-c(curr[1],curr[2])+c(-1,0)
    ### push all of the directions
    stk$push(u)
    stk$push(d)
    stk$push(r)
    stk$push(l)}
  ### return the dictionary with the new basin
  bt}
```


reset the heightmap & create the dictionary

```{r}
heightmap<-matrix(0,ncol=nchar(input[1]),nrow=length(input))
for(i in 1:length(input)){heightmap[i,]<-as.numeric(unlist(str_split(input[i],"")))}

```


Find all of the 9s, as they're not in the basins. 
```{r}
beenthere<-dict()
nines<-which(heightmap==9,arr.ind = TRUE)
for(i in 1:nrow(nines)){
  beenthere$set(c(unname(nines[i,2]),unname(nines[i,1])),"nines")}
```

Then run through every point -
if it's already in the nines or a basin, move to the next.  Otherwise, run it and add to the dictionary

```{r}
i<-0
while(i<=ncol(heightmap)){
  i<-i+1
  j<-0
  while(j<=nrow(heightmap)){
    j<-j+1
    ### cycle through each of these, until everything is filtered into a basin
    if(beenthere$has(c(i,j))){next}
    beenthere<-basins(i,j,beenthere,heightmap)}}

beenthere$size()
```
```{r}
### Group the basins
part2<-unlist(beenthere$values())
### remove the nines
part2<-part2[which(part2!="nines")]
part2<-sort(table(part2),decreasing=TRUE)
part2[1:3]
prod(part2[1:3])

### 95,94,92 (821560)
```

```{r,echo=FALSE}
hmg<-melt(heightmap[2:(nrow(heightmap)-1),2:(ncol(heightmap)-1)])

basicgraph<-ggplot(data=hmg,aes(x=Var2,y=Var1,fill=value))+
  geom_tile()+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position="none")+
  scale_y_reverse()+
  coord_fixed()
basicgraph
```

