---
title: "Day 4 Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
library(knitr)
library(dplyr)
library(stringr)
library(tidyverse)
library(readr)
library(collections)
options(scipen = 999)
```

```{r}
input<-read_lines("Day4Sample.txt")

```

## Part 1
I think this is all about setup - probably.

First, pull the called numbers

```{r}
called<-as.numeric(unlist(str_split(input[1],",")))
input<-input[-c(1,2)]
input<-str_trim(str_squish(input))
```

Create a list of matrices
```{r}

bingocards<-lapply(seq(from=1,to=length(input),by=6),function(x){
  m<-matrix(0,ncol=5,nrow=5)
  for(j in 0:4){
    y<-as.numeric(unlist(str_split(input[x+j]," ")))
    m[j+1,]<-y}
  m})


```

Then, for each card, figure out when each piece is going to get called:
```{r}
bingocalled<-lapply(bingocards,function(x){
  ### create a 5x5 matrix
  bc<-matrix(0,ncol=5,nrow=5)
  ### put the value in where it is called
  bc[]<-vapply(x,function(y){which(called==y)},numeric(1))
  bc})
```
next, figure out when each card will win:

```{r}
bingowin<-sapply(bingocalled,function(x){
  ### not a generic function, because bingo cards are 5x5
  ### find the max of each row & column, row or column could win
  w<-c(apply(x,2,max),apply(x,1,max))
  ### the card will win when the first row or column could win
  w<-min(w)
  w})
```

then (for each card) figure out the sum of the unmarked numbers when it wins

```{r}
unmarked<-sapply(1:length(bingocards),function(x){
  card<-bingocards[[x]]
  called<-bingocalled[[x]]
  win<-bingowin[[x]]
  sc<-matrix(0,ncol=5,nrow=5)
  for(i in 1:5){
    for(j in 1:5){
      sc[i,j]<-if(called[i,j]<=win){0}else{card[i,j]}}}
  sc<-sum(sc)
  sc})
  
```

finally, the score for each is number called * the sum of the unmarked numbers

```{r}
bingoscore<-called[bingowin]*unmarked
bingoscore
### and the winner is the score which corresponds to the lowest win
part1<-bingoscore[which(bingowin==min(bingowin))]
part1
```

## Part 2
So, because we know the scores for all the boards and the order in which they will win, we can just go straight to part 2

```{r}
part2<-bingoscore[which(bingowin==max(bingowin))]
part2
```

